---
title: "Sentiment_analysis"
author: "Jess Chen  |"
date: "April 19, 2026"
format: gfm
editor: visual
---

## Introduction

In this project, we used tidyverse and various sentiment lexicons to conduct sentiment analysis. The first part of this document reproduces sentiment analysis from the book "Text Mining with R" and the latter leverages LLM to create a real time sentiment analysis of FANG companies using Loughran lexicon.

## Part A: Data Reproduction

Our first objective is to reproduce the materials in Chapter 2 of Text Mining with R (Julia Silge and David Robinson). From this exercise, we learn the basic structure of conducting sentiment analysis using Jane Austen novels as an example.

First, we load sentiments into our file.

```{r}

library(tidytext)

get_sentiments("afinn")
get_sentiments("bing")
get_sentiments("nrc")


```

Next, we want to un-nest the words in each novel, then limit the data to 80 lines and analyze each word's sentiment using bing, ultimately resulting in a final sentiment score per novel.

```{r}

library(janeaustenr)
library(dplyr)
library(stringr)
library(tidyr)

#clean data, unnest words
tidy_books <- austen_books() %>%
  group_by(book) %>%
  mutate(
    linenumber = row_number(),
    chapter = cumsum(str_detect(text, 
                                regex("^chapter [\\divxlc]", 
                                      ignore_case = TRUE)))) %>%
  ungroup() %>%
  unnest_tokens(word, text)

#limit to 80 lines, create separate columns and calculate sentiment
jane_austen_sentiment <- tidy_books %>%
  inner_join(get_sentiments("bing")) %>%
  count(book, index = linenumber %/% 80, sentiment) %>%
  pivot_wider(names_from = sentiment, values_from = n, values_fill = 0) %>% 
  mutate(sentiment = positive - negative)

#plot
library(ggplot2)

ggplot(jane_austen_sentiment, aes(index, sentiment, fill = book)) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~book, ncol = 2, scales = "free_x")
```

We also want to compare how the different lexicon libraries fair in terms of analysis.

```{r}

pride_prejudice <- tidy_books %>% 
  filter(book == "Pride & Prejudice")

pride_prejudice

afinn <- pride_prejudice %>% 
  inner_join(get_sentiments("afinn")) %>% 
  group_by(index = linenumber %/% 80) %>% 
  summarise(sentiment = sum(value)) %>% 
  mutate(method = "AFINN")

bing_and_nrc <- bind_rows(
  pride_prejudice %>% 
    inner_join(get_sentiments("bing")) %>%
    mutate(method = "Bing et al."),
  pride_prejudice %>% 
    inner_join(get_sentiments("nrc") %>% 
                 filter(sentiment %in% c("positive", 
                                         "negative"))
    ) %>%
    mutate(method = "NRC")) %>%
  count(method, index = linenumber %/% 80, sentiment) %>%
  pivot_wider(names_from = sentiment,
              values_from = n,
              values_fill = 0) %>% 
  mutate(sentiment = positive - negative)

bind_rows(afinn, 
          bing_and_nrc) %>%
  ggplot(aes(index, sentiment, fill = method)) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~method, ncol = 1, scales = "free_y")

```

It's also relevant to look beyond words, at sentences because words taken in isolation can sometimes be misleading.

```{r}
#split text into dataframes by chapters

austen_chapters <- austen_books() %>%
  group_by(book) %>%
  unnest_tokens(chapter, text, token = "regex", 
                pattern = "Chapter|CHAPTER [\\dIVXLC]") %>%
  ungroup()

austen_chapters %>% 
  group_by(book) %>% 
  summarise(chapters = n())

#Find negative words by chapter

bingnegative <- get_sentiments("bing") %>% 
  filter(sentiment == "negative")

wordcounts <- tidy_books %>%
  group_by(book, chapter) %>%
  summarize(words = n())

tidy_books %>%
  semi_join(bingnegative) %>%
  group_by(book, chapter) %>%
  summarize(negativewords = n()) %>%
  left_join(wordcounts, by = c("book", "chapter")) %>%
  mutate(ratio = negativewords/words) %>%
  filter(chapter != 0) %>%
  slice_max(ratio, n = 1) %>% 
  ungroup()



```

## Part B: Additional corpus

I would like to use sentiment analysis to look at news headlines and assess public sentiment against companies (for stock/portfolio analysis. To simplify the task, I will use FANG companies.

```{r}
library(tidyquant)
library(purrr)
library(rvest)
library(dplyr)
library(tidytext)
library(tidyverse)

# Function to scrape headlines for any ticker
get_ticker_news <- function(symbol) {
  url <- paste0("https://finviz.com/quote.ashx?t=", symbol)
  
  # Read the page
  webpage <- read_html(url)
  
  # Extract the news table using its specific ID
  news_table <- webpage %>%
    html_nodes(xpath = '//*[@id="news-table"]') %>%
    html_table() %>%
    .[[1]] %>%
    setNames(c("time", "headline")) %>%
    mutate(ticker = symbol) # Keep track of which company is which
  
  return(news_table)
}

#Run for mulitple companies

# Define your FANG / MAMAA list
tickers <- c("META", "AMZN", "NFLX", "GOOGL", "AAPL", "MSFT", "NVDA")

# Map the function across all tickers and combine into one big table
all_news <- map_df(tickers, get_ticker_news)

# Quick look at what we got
head(all_news)


#Run analysis
# Tokenize the headlines
tidy_news <- all_news %>%
  unnest_tokens(word, headline)

# Join with Loughran lexicon
# Note: You may be prompted to install 'textdata' or agree to terms in the console
sentiment_counts <- tidy_news %>%
  inner_join(get_sentiments("loughran")) %>%
  filter(sentiment %in% c("positive", "negative")) %>%
  count(ticker, sentiment) %>%
  spread(sentiment, n, fill = 0) %>%
  mutate(score = positive - negative)


#Visualize

ggplot(sentiment_counts, aes(x = reorder(ticker, score), y = score, fill = score > 0)) +
  geom_col() +
  coord_flip() +
  labs(title = "FANG / Tech Sentiment: Current News Headlines",
       subtitle = "Using Loughran-McDonald Financial Lexicon",
       x = "Company Ticker",
       y = "Net Sentiment Score (Pos - Neg)") +
  theme_minimal() +
  guides(fill = "none")
```

## Conclusion

I find sentiment analysis really interesting. In today's highly connected and social world, popular internet sentiment can prove pivotal to understanding & predicting market behaviour. In the second part of the analysis, it is interesting to see the market's shift away from NVDA, once a market darling, into more traditional companies like Google and Apple.

## References:

Gemini Fast (2026) Google LLM. Link to chat: <https://gemini.google.com/share/96f36b0f8665>

Silge, Julia, and David Robinson. *Text Mining with R: A Tidy Approach*. Accessed \[April 19\]. <https://www.tidytextmining.com/>.
